Miniproject - Analytic Number Theory

### 2021-12-15 Analytic number theory/submission.tex
small fixed, mainly to ANT

    To show surjectivity, consider some Dirichlet character $\chi: \N \to \C$ and construct a group homomorphism $f: (\Z/q\Z)^\units \to \C^\units$. 
    For each $x \in (\Z/q\Z)^\units$, there is a representative $n \in \N$ of $x$ and as $\chi(n)$ does not depend on the choice of $n$, we may define $f(x) := \chi(n)$.
    Note that as $x \in (\Z/q\Z)^\units$, we find $n \perp q$ and so $\chi(n) \neq 0$, i.e. $\chi(n) \in \C^*$.
    Then clearly for $a, b \in (\Z/q\Z)^*$ with representatives $n, m \in \N$ have
    \begin{equation*}
        f(ab) = \chi(nm) = \chi(n)\chi(m) = f(a)f(b)
    \end{equation*}
    So $f$ is a well-defined group homomorphism and we obviously have $\chi_f = \chi$.
\end{proof}
For simplicity of notation we sometimes will identify a Dirichlet character and its group homomorphism if it is always clear which one is meant.
\begin{example}[Ex (i)]
    \label{ex:nontrivial_dirichlet_character_mod_4}
    The function
    \begin{equation*}
        f: \N \to \C, \quad n \mapsto \begin{cases}
            0 & \text{if $n \equiv 0, 2 \mod 4$} \\
            1 & \text{if $n \equiv 1 \mod 4$} \\
            -1 & \text{if $n \equiv 3 \mod 4$}
        \end{cases}
    \end{equation*}
    is a Dirichlet character.
\end{example}
\begin{proof}
    This follows directly from Lemma~\ref{prop:characterization_dirichlet_character}, as $f = \chi_g$ for the group homomorphism
    \begin{equation*}
        g: (\Z/4\Z)^\units = \{ 1, 3 \} \to \C^*, \quad 1 \mapsto 1, \ 3 \mapsto -1
    \end{equation*}
    (this is a group homomorphism, as $3^2 = 9 \equiv 1 \mod 4$)
\end{proof}
Now we want to define Dirichlet series of Dirichlet characters.
\begin{proposition}
    For a Dirichlet character $\chi: \N \to \C$ and some $\epsilon > 0$, the series
    \begin{equation*}
        L(s, \chi) := \sum_{n \geq 1} \chi(n) n^{-s}
    \end{equation*}
    converges uniformly on $\Re(s) \geq 1 + \epsilon$.
    We will call it the Dirichlet series of $\chi$.
\end{proposition}
\begin{proof}
    By Lemma~\ref{prop:characterization_dirichlet_character}, we know that $\chi$ corresponds to a group homomorphism $f: (\Z/q\Z)^\units \to \C^\units$ such that $\chi(\N) = f((\Z/q\Z)^*) \cup \{ 0 \} \subseteq \C$ is a finite subset of $\C$.
    Hence, there is $C > 0$ with $|\chi(n)| \leq C$ for all $n \in \N$, and it follows that
    \begin{equation*}
        \sum_{1 \leq n \leq X} \left| \chi(n) n^{-s} \right| \leq \sum_{1 \leq n \leq X} C \left| n^{-s} \right| \leq C \sum_{1 \leq n \leq X} n^{-1 - \epsilon} \leq C \sum_{n \geq 1} n^{-1 - \epsilon}
    \end{equation*}
    which is finite.
\end{proof}
\begin{proposition}
    \label{prop:dirichlet_series_at_one}
    Let $f: (\Z/q\Z)^\units \to \C^\units$ be a group homomorphism. Then for the associated Dirichlet character $\chi = \chi_f$ we have that
    \begin{equation*}
        \lim_{s \to 1^+} L(s, \chi) \ \text{exists} \ \Leftrightarrow \ \sum_{x \in (\Z/q\Z)^\units} f(x) = 0
    \end{equation*}
    If this is the case, then
    \begin{equation*}
        \lim_{s \to 1^+} L(s, \chi) = \sum_{n \geq 1} \frac {\chi(n)} n
    \end{equation*}
    where the right-hand side converges (but not absolutely).
\end{proposition}
\begin{proof}
    Let $c = \sum_{x \in (\Z/q\Z)^\units} f(x)$.
    For the direction $\Rightarrow$ assume that $c \neq 0$.
    Then have for $\Re(s) > 1$ that
    \begin{align*}
        \sgn(c) \sum_{n \geq 1} \chi(n) n^{-s} &= \sum_{n \geq 0} \ \sum_{0 < k \leq q} \sgn(c) \chi(qn + k) (qn + k)^{-s} \\
        &\geq \sum_{n \geq 0} \ \sum_{0 < k \leq q}\sgn(c) \chi(qn + k) (qn + n)^{-s} \\
        &= \sum_{n \geq 0} \sgn(c) (qn + n)^{-s} \underbrace{\sum_{0 < k \leq q} \chi(qn + k)}_{= c} \\
        &\geq \frac {|c|} {(q + 1)^s} \sum_{n \geq 1} n^{-s} = \frac {|c|} {(q + 1)^s} \zeta(s)
    \end{align*}
    which clearly has a pole at $s = 1$. Hence $\lim_{s \to 1^+} L(s, \chi_f)$ cannot exist.

@@

        =& \ O(qCX^{-s}) + O\Bigl( Cqs \sum_{X/q \leq n} n^{-(1 + \epsilon)} \Bigr)

@@

\begin{example}[Ex (ii)]

@@

\begin{corollary}[Ex (iii)]

@@

For the case of primes $\equiv 1 \mod 4$, I have remembered the two-square theorem and its connection to primes in the ring $\Z[i]$ of Gaussian integers, and somehow my train of thoughts went into Algebraic Number Theory.

@@

    Now consider the element $\alpha = \omega_q - xqP \in \mathcal{O}$. Then
    \begin{align*}
        N_{K/\Q}(\alpha) &= \prod_{\sigma: K \to \C \ \text{$\Q$-field homomorphism}} \sigma(\omega_q - xqP) \\
        &= \prod_{\sigma} (\sigma(\omega_q) - xqP) = \mathrm{MiPo}_\Q(\omega_q)(xqP) = \Phi_q(xqP) \neq 1
    \end{align*}
    as $\mathrm{MiPo}_\Q(\omega_q) = \prod_\sigma (\sigma(\omega_q) - X)$.
    Hence, $\alpha$ is not a unit in $\mathcal{O}$.
    On the other hand, $(\alpha)$ is coprime to $(p_i)$ for each $p_i$, as 
    \begin{equation*}
        \omega_q = \alpha - xqP \in (\alpha) + (p_i) \quad \text{and} \quad \omega_q \in \mathcal{O}^\units
    \end{equation*}
    By our assumption, the only prime ideals in $\mathcal{O}$ are the prime ideal factors of $(p_i)$ and $(p)$ for $p \neq p_i$.
    Thus, the prime ideal factorization of $(\alpha)$ consists only of prime ideals $(p), p \neq p_i$ and it follows that $(\alpha) = (n)$ for some integer $n \geq 2$.
    As $\omega_q$ and $xqP \in \Z$ are $\Q$-linearly independent, we see that $n \divides \omega_q$ and $n \divides xqP$.
    However, the former is a contradiction, as $\omega_q \in \mathcal{O}^\units$ is a unit and no $n \geq 2$ is a unit.
\end{proof}
The book also mentions that the general case can be proven by using L-series in algebraic number fields.
\begin{corollary}[Ex (iii)]
    There are infinitely many primes $p$ with $p \equiv 1 \mod 4$.
\end{corollary}
\begin{proof}
    This is just a special case of Prop.~\ref{prop:primes_1_mod_q}.
\end{proof}
\begin{example}[Ex (iii)]
    Using a computer, we can also study the actual frequency of prime numbers $\equiv 1, 3 \mod 4$ among e.g. the first $10^8$ integers.
    This seems to indicate that both numbers are asymptotically equal, which seems natural, given the result of Example~\ref{ex:dirichlet_sum_over_primes}.
    For example, there are 332180 primes $\equiv 1 \mod 4$ and 332398 primes $\equiv 3 \mod 4$ smaller than $10^8$.
    To find these numbers, the following python code was used.
    \lstinputlisting[language = python]{./primes.py}
\end{example}

\section{Part II}
We have already shown that Dirichlet characters are, in principle, group homomorphisms $(\Z/q\Z)^\units \to \C^\units$.
If we now assume $q$ to be prime, we get an even nicer characterization.
So for the whole section, assume that $q \geq 3$ is a prime.
\begin{corollary}[Ex (i)]
    \label{prop:dirichlet_character_image_group}
    Let $\chi, \chi': \N \to \C$ be Dirichlet characters mod $q$ and $r$ a representative of a primitive root modulo $q$.
    If $\chi(r) = \chi'(r)$, then $\chi = \chi'$.
    Further, have that $\chi(n)^{q - 1} = 1$ for all $n \in \N$ with $n \perp q$.
\end{corollary}
\begin{proof}
    The properties follow directly from Lemma~\ref{prop:characterization_dirichlet_character}.
    Let $f, f': (\Z/q\Z)^\units \to \C^\units$ be the associated group homomorphisms of $\chi, \chi'$ as in Lemma~\ref{prop:characterization_dirichlet_character}.
    If $f([r]) = \chi(r) = \chi'(r) = f'([r])$ then clearly $f = f'$, as these are group homomorphisms and $\langle [r] \rangle = (\Z/q\Z)^\units$.
    Hence $\chi = \chi'$.

    Further, have for $n \in \N$ with $n \perp q$ that $[n] \in (\Z/q\Z)^\units$ and thus
    \begin{equation*}
        [n]^{q - 1} = [n]^{\phi(q)} = [n]^{|(\Z/q\Z)^\units|} = 1 
    \end{equation*}
    As $f$ is a group homomorphism, find
    \begin{equation*}
        \chi(n)^{q - 1} = f([n])^{q - 1} = f([n]^{q - 1}) = f(1) = 1
    \end{equation*}
\end{proof}
This correspondence also works in the other direction.
\begin{corollary}[Ex (ii)]
    \label{prop:dirichlet_character_primitive_root}
    Let $\omega \in \C$ be a $(q - 1)$-th root of unity, i.e. $\omega^{q - 1} = 1$ and let $r \in \Z$ be a representative of a primitive root modulo $q$.
    Then
    \begin{equation*}
        g: \N \to \C, \quad n \mapsto \begin{cases}
            \omega^{\log_r n} & \text{if $n \perp q$} \\
            0 & \text{otherwise}
        \end{cases}
    \end{equation*}
    is a well-defined Dirichlet character.
\end{corollary}
\begin{proof}
    Follows again directly from Lemma~\ref{prop:characterization_dirichlet_character}, as $[r] \mapsto \omega$ induces a unique group homomorphism $(\Z/q\Z)^\units \to \C^\units$.
    The associated Dirichlet character is obviously $g$.
\end{proof}
Note that the image of a group homomorphism $f: (\Z/q\Z)^\units \to \C^\units$ is a subgroup of $\C^\units$.
Using Corollary~\ref{prop:dirichlet_character_image_group}, we can describe it quite concretely.
\begin{proposition}
    Let $\chi: \N \to \C$ be a Dirichlet character with group homomorphism $f: (\Z/q\Z)^\units \to \C^\units$.
    Then $\mathrm{im}f \leq S$ is a subgroup where $S_q := \{ \omega_q^k \ | \ k \in \Z \}$ is the group of $q$-th roots of unity.

    It is a fact from Algebra that $S_q \cong (\Z/q\Z)^\units$, hence Dirichlet characters modulo a prime $q$ are in 1-to-1 correspondence with the endomorphisms $\mathrm{End}((\Z/q\Z)^\units)$ of $(\Z/q\Z)^\units$.
\end{proposition}
\begin{proof}
    We have that $S_q = \{ x \in \C^\units \ | \ x^{q - 1} = 1 \}$ and the claim directly follows from Corollary~\ref{prop:dirichlet_character_image_group}.
\end{proof}
Note that the endomorphism monoid $\mathrm{End}((\Z/q\Z)^\units)$ is not a group, except in the trivial case $q = 2$.
The reason is that e.g. the trivial group homomorphism $r \mapsto 1$ is not surjective and thus not invertible.
\begin{definition}
    Denote by $\mathrm{Dir}(q)$ the set of Dirichlet characters modulo $q$.
\end{definition}
By Corollary~\ref{prop:dirichlet_character_primitive_root} each group endomorphism $f \in \mathrm{End}((\Z/q\Z)^\units)$ is determined by its value at a primitive root of unity $r \in (\Z/q\Z)^\units$, hence
\begin{equation*}
    |\mathrm{Dir}(q)| = |\mathrm{End}((\Z/q\Z)^\units)| = |(\Z/q\Z)^\units| = q - 1
\end{equation*}
It follows that there are exactly $q - 1$ distinct Dirichlet characters modulo a prime $q$.
\begin{remark}
    It is again a fact that $(\Z/p^k\Z)^\units$ is cyclic for an odd prime $p$ and $k \geq 1$.
    Hence, everything up to now can also be done for odd prime powers, if we replace $q - 1$ by $\phi(q)$.
\end{remark}
Because of Lemma~\ref{prop:dirichlet_series_at_one} it might seem like a good idea to study in which cases the value $\sum_{x \in (\Z/q\Z)^\units} \chi(x)$ is zero.
\begin{proposition}[Ex (iii)]
    \label{prop:finite_sums_characters}
    Let $\chi_0$ be the trivial Dirichlet character given by $r \mapsto 1$.
    Then
    \begin{align*}
        \sum_{a \in (\Z/q\Z)^\units} \chi(a) &= \begin{cases}
            q - 1 & \text{if $\chi = \chi_0$} \\
            0 & \text{otherwise}
        \end{cases}, \\
        \sum_{\chi \in \mathrm{Dir}(q)} \chi(a) &= \begin{cases}
            q - 1 & \text{if $a \equiv 1 \mod q$} \\
            0 & \text{otherwise}
        \end{cases}
    \end{align*}
    Furthermore, for $b \perp q$ have
    \begin{equation*}
        \sum_{\chi \in \mathrm{Dir}(q)} \chi(a)\overline{\chi(b)} = \begin{cases}
            q - 1 & \text{if $a \equiv b \mod q$} \\
            0 & \text{otherwise}
        \end{cases}
    \end{equation*}
\end{proposition}
\begin{proof}
    Clearly
    \begin{equation*}
        \sum_{a \in (\Z/q\Z)^\units} \chi_0(a) = q - 1 \quad \text{and} \quad \sum_{\chi \in \mathrm{Dir}(q)} \chi(1) = \sum_{\chi \in \mathrm{Dir}(q)} 1 = q - 1
    \end{equation*}
    So it is left to show that we get zero in the other cases.

    Consider a Dirichlet character $\chi \neq \chi_0$ given by $r \mapsto \xi$ for a $q$-th root of unity $\xi \neq 1$.
    Then
    \begin{equation*}
        \sum_{a \in (\Z/q\Z)^\units} \chi(a) = \sum_{k = 0}^{q - 2} \chi(r^k) = \sum_{k = 0}^{q - 2} \xi^k = \frac {1 - \xi^{q - 1}} {q - \xi} = 0 
    \end{equation*}
    By using the earlier results on the structure of $\mathrm{Dir}(q)$ we see that for $a \equiv r^k \not\equiv 1 \mod q$, have
    \begin{align*}
        \sum_{\chi \in \mathrm{Dir}(q)} \chi(a) &= \sum_{\chi \in \mathrm{Dir}(q)} \chi(r)^k \\
        &= \sum_{\xi \ \text{$q$-th root of unity}} \xi^k = \sum_{l = 0}^{q - 2} \omega^{kl} = \frac {1 - (\omega^{q - 1})^k} {1 - \omega^k} = 0 
    \end{align*}
    where $\omega \in \C$ is a primitive $q$-th root of unity and $r \in \Z$ is a primitive root modulo $q$.

>>>>

    To show surjectivity, consider some Dirichlet character $\chi: \N \to \C$ and construct a group homomorphism $f: (\Z/q\Z)^\units \to \C^\units$. 
    For each $x \in (\Z/q\Z)^\units$, there is a representative $n \in \N$ of $x$ and as $\chi(n)$ does not depend on the choice of $n$, we may define $f(x) := \chi(n)$.
    Note that as $x \in (\Z/q\Z)^\units$, we find $n \perp q$ and so $\chi(n) \neq 0$, i.e. $\chi(n) \in \C^*$.
    Then clearly for $a, b \in (\Z/q\Z)^*$ with representatives $n, m \in \N$ have
        f(ab) = \chi(nm) = \chi(n)\chi(m) = f(a)f(b)
    So $f$ is a well-defined group homomorphism and we obviously have $\chi_f = \chi$.
For simplicity of notation we sometimes will identify a Dirichlet character and its group homomorphism if it is always clear which one is meant.
\begin{example}[Question (i)]
    \label{ex:nontrivial_dirichlet_character_mod_4}
    The function
        f: \N \to \C, \quad n \mapsto \begin{cases}
            0 & \text{if $n \equiv 0, 2 \mod 4$} \\
            1 & \text{if $n \equiv 1 \mod 4$} \\
            -1 & \text{if $n \equiv 3 \mod 4$}
        \end{cases}
    is a Dirichlet character.
    This follows directly from Lemma~\ref{prop:characterization_dirichlet_character}, as $f = \chi_g$ for the group homomorphism
        g: (\Z/4\Z)^\units = \{ 1, 3 \} \to \C^*, \quad 1 \mapsto 1, \ 3 \mapsto -1
    (this is a group homomorphism, as $3^2 = 9 \equiv 1 \mod 4$)
Now we want to define Dirichlet series of Dirichlet characters.
    For a Dirichlet character $\chi: \N \to \C$ and some $\epsilon > 0$, the series
        L(s, \chi) := \sum_{n \geq 1} \chi(n) n^{-s}
    converges uniformly on $\Re(s) \geq 1 + \epsilon$.
    We will call it the Dirichlet series of $\chi$.
    By Lemma~\ref{prop:characterization_dirichlet_character}, we know that $\chi$ corresponds to a group homomorphism
        f: (\Z/q\Z)^\units \to \C^\units
    such that $\chi(\N) = f((\Z/q\Z)^*) \cup \{ 0 \} \subseteq \C$ is a finite subset of $\C$.
    Hence, there is $C > 0$ with $|\chi(n)| \leq C$ for all $n \in \N$, and it follows that
        \sum_{1 \leq n \leq X} \left| \chi(n) n^{-s} \right| \leq \sum_{1 \leq n \leq X} C \left| n^{-s} \right| \leq C \sum_{1 \leq n \leq X} \left| n^{-s} \right|
    We know that the series
        \sum_{n \geq 1} n^{-s}
    converges uniformly on $\Re(s) \geq 1 + \epsilon$ and so the claim follows.
    \label{prop:dirichlet_series_at_one}
    Let $f: (\Z/q\Z)^\units \to \C^\units$ be a group homomorphism. Then for the associated Dirichlet character $\chi = \chi_f$ we have that
        \lim_{s \to 1^+} L(s, \chi) \ \text{exists} \ \Leftrightarrow \ \sum_{x \in (\Z/q\Z)^\units} f(x) = 0
    If this is the case, then
        \lim_{s \to 1^+} L(s, \chi) = \sum_{n \geq 1} \frac {\chi(n)} n
    where the right-hand side converges (but not absolutely).
    Let $c = \sum_{x \in (\Z/q\Z)^\units} f(x)$.
    For the direction $\Rightarrow$ assume that $c \neq 0$.
    Then have for $\Re(s) > 1$ that
        \sgn(c) \sum_{n \geq 1} \chi(n) n^{-s} &= \sum_{n \geq 0} \ \sum_{0 < k \leq q} \sgn(c) \chi(qn + k) (qn + k)^{-s} \\
        &\geq \sum_{n \geq 0} \ \sum_{0 < k \leq q}\sgn(c) \chi(qn + k) (qn + n)^{-s} \\
        &= \sum_{n \geq 0} \sgn(c) (qn + n)^{-s} \underbrace{\sum_{0 < k \leq q} \chi(qn + k)}_{= c} \\
        &\geq \frac {|c|} {(q + 1)^s} \sum_{n \geq 1} n^{-s} = \frac {|c|} {(q + 1)^s} \zeta(s)
    which clearly has a pole at $s = 1$. Hence $\lim_{s \to 1^+} L(s, \chi_f)$ cannot exist.

@@

        =& \ O(qCX^{-s}) + O\Bigl( Cqs \sum_{n \geq X/q} n^{-(1 + \epsilon)} \Bigr)

@@

\begin{example}[Question (ii)]

@@

\begin{corollary}[Question (iii)]

@@

For the case of primes $\equiv 1 \mod 4$, I have remembered Fermat's theorem on the sum of two squares and its connection to primes in the ring $\Z[i]$ of Gaussian integers, and somehow my train of thoughts went into Algebraic Number Theory.

@@

    Now consider the element $\alpha = \omega_q - xqP \in \mathcal{O}$. Then
    \begin{align*}
        N_{K/\Q}(\alpha) &= \prod_{\sigma: K \to \C \ \text{$\Q$-field homomorphism}} \sigma(\omega_q - xqP) \\
        &= \prod_{\sigma} (\sigma(\omega_q) - xqP) = \mathrm{MiPo}_\Q(\omega_q)(xqP) = \Phi_q(xqP) \neq 1
    \end{align*}
    as $\mathrm{MiPo}_\Q(\omega_q) = \prod_\sigma (\sigma(\omega_q) - X)$.
    Hence, $\alpha$ is not a unit in $\mathcal{O}$.
    On the other hand, $(\alpha)$ is coprime to $(p_i)$ for each $p_i$, as 
    \begin{equation*}
        \omega_q = \alpha - xqP \in (\alpha) + (p_i) \quad \text{and} \quad \omega_q \in \mathcal{O}^\units
    \end{equation*}
    By our assumption, the only prime ideals in $\mathcal{O}$ are the prime ideal factors of $(p_i)$ and $(p)$ for $p \neq p_i$.
    Thus, the prime ideal factorization of $(\alpha)$ consists only of prime ideals $(p), p \neq p_i$ and it follows that $(\alpha) = (n)$ for some integer $n \geq 2$.
    As $\omega_q$ and $xqP \in \Z$ are $\Q$-linearly independent, we see that $n \divides \omega_q$ and $n \divides xqP$.
    However, the former is a contradiction, as $\omega_q \in \mathcal{O}^\units$ is a unit and no $n \geq 2$ is a unit.
\end{proof}
The book also mentions that the general case can be proven by using L-series in algebraic number fields.
I am pretty sure that those are exactly the functions $L(\chi, s)$ that we defined above.
\begin{corollary}[Question (iii)]
    There are infinitely many primes $p$ with $p \equiv 1 \mod 4$.
\end{corollary}
\begin{proof}
    This is just a special case of Prop.~\ref{prop:primes_1_mod_q}.
\end{proof}
\begin{example}[Question (iii)]
    Using a computer, we can also study the actual frequency of prime numbers $\equiv 1, 3 \mod 4$ among e.g. the first $10^8$ integers.
    This seems to indicate that both numbers are asymptotically equal, which seems natural, given the result of Example~\ref{ex:dirichlet_sum_over_primes}.
    For example, there are 332180 primes $\equiv 1 \mod 4$ and 332398 primes $\equiv 3 \mod 4$ smaller than $10^8$.
    To find these numbers, the following python code was used.
    \lstinputlisting[language = python]{./primes.py}
\end{example}

\section{Part II}
We have already shown that Dirichlet characters are, in principle, group homomorphisms $(\Z/q\Z)^\units \to \C^\units$.
If we now assume $q$ to be prime, we get an even nicer characterization.
So for the whole section, assume that $q \geq 3$ is a prime.
\begin{corollary}[Question (i)]
    \label{prop:dirichlet_character_image_group}
    Let $\chi, \chi': \N \to \C$ be Dirichlet characters mod $q$ and $r$ a representative of a primitive root modulo $q$.
    If $\chi(r) = \chi'(r)$, then $\chi = \chi'$.
    Further, have that $\chi(n)^{q - 1} = 1$ for all $n \in \N$ with $n \perp q$.
\end{corollary}
\begin{proof}
    The properties follow directly from Lemma~\ref{prop:characterization_dirichlet_character}.
    Let $f, f': (\Z/q\Z)^\units \to \C^\units$ be the associated group homomorphisms of $\chi, \chi'$ as in Lemma~\ref{prop:characterization_dirichlet_character}.
    If $f([r]) = \chi(r) = \chi'(r) = f'([r])$ then clearly $f = f'$, as these are group homomorphisms and $\langle [r] \rangle = (\Z/q\Z)^\units$.
    Hence $\chi = \chi'$.

    Further, have for $n \in \N$ with $n \perp q$ that $[n] \in (\Z/q\Z)^\units$ and thus
    \begin{equation*}
        [n]^{q - 1} = [n]^{\phi(q)} = [n]^{|(\Z/q\Z)^\units|} = 1 
    \end{equation*}
    As $f$ is a group homomorphism, find
    \begin{equation*}
        \chi(n)^{q - 1} = f([n])^{q - 1} = f([n]^{q - 1}) = f(1) = 1
    \end{equation*}
\end{proof}
This correspondence also works in the other direction.
\begin{corollary}[Question (ii)]
    \label{prop:dirichlet_character_primitive_root}
    Let $\omega \in \C$ be a $(q - 1)$-th root of unity, i.e. $\omega^{q - 1} = 1$ and let $r \in \Z$ be a representative of a primitive root modulo $q$.
    Then
    \begin{equation*}
        g: \N \to \C, \quad n \mapsto \begin{cases}
            \omega^{\log_r n} & \text{if $n \perp q$} \\
            0 & \text{otherwise}
        \end{cases}
    \end{equation*}
    is a well-defined Dirichlet character.
\end{corollary}
\begin{proof}
    Follows again directly from Lemma~\ref{prop:characterization_dirichlet_character}, as $[r] \mapsto \omega$ induces a unique group homomorphism $(\Z/q\Z)^\units \to \C^\units$.
    The associated Dirichlet character is obviously $g$.
\end{proof}
Note that the image of a group homomorphism $f: (\Z/q\Z)^\units \to \C^\units$ is a subgroup of $\C^\units$.
Using Corollary~\ref{prop:dirichlet_character_image_group}, we can describe it quite concretely.
\begin{proposition}
    Let $\chi: \N \to \C$ be a Dirichlet character with group homomorphism $f: (\Z/q\Z)^\units \to \C^\units$.
    Then $\mathrm{im}f \leq S$ is a subgroup where $S_q := \{ \omega_q^k \ | \ k \in \Z \}$ is the group of $q$-th roots of unity.

    It is a fact from Algebra that $S_q \cong (\Z/q\Z)^\units$, hence Dirichlet characters modulo a prime $q$ are in 1-to-1 correspondence with the endomorphisms $\mathrm{End}((\Z/q\Z)^\units)$ of $(\Z/q\Z)^\units$.
\end{proposition}
\begin{proof}
    We have that $S_q = \{ x \in \C^\units \ | \ x^{q - 1} = 1 \}$ and the claim directly follows from Corollary~\ref{prop:dirichlet_character_image_group}.
\end{proof}
Note that the endomorphisms $\mathrm{End}((\Z/q\Z)^\units)$ form a ring w.r.t. pointwise multiplication and composition
(this is the case for the endomorphisms of any abelian group).
In particular, the endomorphisms and so the Dirichlet characters form a group w.r.t. pointwise multiplication.
\begin{definition}
    Denote by $\mathrm{Dir}(q)$ the set of Dirichlet characters modulo $q$.
\end{definition}
By Corollary~\ref{prop:dirichlet_character_image_group} each Dirichlet character resp. group endomorphism of $(\Z/q\Z)^\units$ is determined by its value at a primitive root of unity $r \in (\Z/q\Z)^\units$.
Further, by Corollary~\ref{prop:dirichlet_character_primitive_root}, we also have a Dirichlet character resp. group endomorphism for each possible value at a primitive root of unity.
Hence
\begin{equation*}
    |\mathrm{Dir}(q)| = |\mathrm{End}((\Z/q\Z)^\units)| = |(\Z/q\Z)^\units| = q - 1
\end{equation*}
It follows that there are exactly $q - 1$ distinct Dirichlet characters modulo a prime $q$.
\begin{remark}
    It is again a fact that $(\Z/p^k\Z)^\units$ is cyclic for an odd prime $p$ and $k \geq 1$.
    Hence, everything up to now can also be done for odd prime powers, if we replace $q - 1$ by $\phi(q)$.
\end{remark}
Because of Lemma~\ref{prop:dirichlet_series_at_one} it might seem like a good idea to study in which cases the value $\sum_{x \in (\Z/q\Z)^\units} \chi(x)$ is zero.
\begin{proposition}[Question (iii)]
    \label{prop:finite_sums_characters}
    Let $\chi_0$ be the trivial Dirichlet character given by $r \mapsto 1$.
    Then
    \begin{align*}
        \sum_{a \in (\Z/q\Z)^\units} \chi(a) &= \begin{cases}
            q - 1 & \text{if $\chi = \chi_0$} \\
            0 & \text{otherwise}
        \end{cases}, \\
        \sum_{\chi \in \mathrm{Dir}(q)} \chi(a) &= \begin{cases}
            q - 1 & \text{if $a \equiv 1 \mod q$} \\
            0 & \text{otherwise}
        \end{cases}
    \end{align*}
    Furthermore, for $b \perp q$ have
    \begin{equation*}
        \sum_{\chi \in \mathrm{Dir}(q)} \chi(a)\overline{\chi(b)} = \begin{cases}
            q - 1 & \text{if $a \equiv b \mod q$} \\
            0 & \text{otherwise}
        \end{cases}
    \end{equation*}
\end{proposition}
\begin{proof}
    Clearly
    \begin{equation*}
        \sum_{a \in (\Z/q\Z)^\units} \chi_0(a) = q - 1 \quad \text{and} \quad \sum_{\chi \in \mathrm{Dir}(q)} \chi(1) = \sum_{\chi \in \mathrm{Dir}(q)} 1 = q - 1
    \end{equation*}
    So it is left to show that we get zero in the other cases.

    Consider a Dirichlet character $\chi \neq \chi_0$ given by $r \mapsto \xi$ for a primitive $q$-th root of unity $\xi \neq 1$.
    Then
    \begin{equation*}
        \sum_{a \in (\Z/q\Z)^\units} \chi(a) = \sum_{k = 0}^{q - 2} \chi(r^k) = \sum_{k = 0}^{q - 2} \xi^k = \frac {1 - \xi^{q - 1}} {q - \xi} = 0 
    \end{equation*}
    By using the earlier results on the structure of $\mathrm{Dir}(q)$ we see that for $a \equiv r^k \not\equiv 1 \mod q$, have
    \begin{align*}
        \sum_{\chi \in \mathrm{Dir}(q)} \chi(a) &= \sum_{\chi \in \mathrm{Dir}(q)} \chi(r)^k \\
        &= \sum_{\xi \ \text{$q$-th root of unity}} \xi^k = \sum_{l = 0}^{q - 2} \omega^{kl} = \frac {1 - (\omega^{q - 1})^k} {1 - \omega^k} = 0 
    \end{align*}
    where $\omega \in \C$ is any primitive $q$-th root of unity and $r \in \Z$ is a primitive root modulo $q$.

    For the last part, note that for any $q$-th root of unity $\xi$, we have $\xi\overline{\xi} \in \R$ with $\xi\overline{\xi} = |\xi|^2 > 0$.
    Furthermore, $\overline{\xi}$ is also a $q$-th root of unity, and so we see that $\xi\overline{\xi} = 1$ (the only real, positive root of unity is 1).
    It follows that for any Dirichlet character $\chi$ have $\overline{\chi([a])} = \chi([a]^{-1})$.
    Thus
    \begin{equation*}
        \sum_{\chi \in \mathrm{Dir}(q)} \chi(a)\overline{\chi(b)} = \sum_{\chi \in \mathrm{Dir}(q)} \chi([a][b]^{-1}) = \begin{cases}
            q - 1 & \text{if $[a][b]^{-1} = 1 \in (\Z/q\Z)^\units$} \\
            0 & \text{otherwise}
        \end{cases}
    \end{equation*}
    The condition $ab^{-1} = 1$ is equivalent to $a \equiv b \mod q$, so the claim follows.
\end{proof}
Using these basic results, we can now prove facts on the Dirichlet series of characters.
\begin{proposition}[Question (iv)]
    \label{prop:representation_mangoldt_series_congruent_a}
    Let $a \perp q$. Then for $\Re(s) > 1$ have
    \begin{equation*}
        \sum_{n \equiv a \mod q} \frac {\Lambda(n)} {n^s} = \frac 1 {q - 1} \sum_{\chi \in \mathrm{Dir}(q)} \overline{\chi(a)} \sum_{n \geq 1} \frac {\Lambda(n) \chi(n)} {n^s}
    \end{equation*}
    (All those series obviously converge absolutely since $\Re(s) > 1$)
\end{proposition}
\begin{proof}
    By Prop.~\ref{prop:finite_sums_characters} we have for all $n \in \N$ that
    \begin{equation*}
        \frac 1 {q - 1} \sum_{\chi \in \mathrm{Dir}(q)} \overline{\chi(a)} \chi(n) = \begin{cases}
            1 & \text{if $a \equiv n \mod q$} \\
            0 & \text{otherwise}
        \end{cases}
    \end{equation*}
    It follows that
    \begin{align*}
        \sum_{n \equiv a \mod q} \Lambda(n) n^{-s} =& \sum_{n \geq 1} \Lambda(n) n^{-s} \begin{cases}
            1 & \text{if $a \equiv n \mod q$} \\
            0 & \text{otherwise}
        \end{cases} \\
        =& \sum_{n \geq 1} \Lambda(n) n^{-s} \frac 1 {q - 1} \sum_{\chi \in \mathrm{Dir}(q)} \overline{\chi(a)} \chi(n) \\
        =& \frac 1 {q - 1} \sum_{\chi \in \mathrm{Dir}(q)} \overline{\chi(a)} \sum_{n \geq 1} \Lambda(n) \chi(n) n^{-s} 
    \end{align*}
    as infinite summation clearly commutes with finite sums.
\end{proof}
\begin{example}[Question (v)]
    \label{ex:nontrivial_dirichlet_characters_mod_5}
    We consider the Dirichlet characters mod 5.
    A (primitive) 5-th root of unity $\omega_5 \in \C$ is given by $\omega_5 = \exp(2\pi i/5)$.
    On the other hand, a primitive root modulo 5 is e.g. given by $r = 2$ since $2^2 \equiv -1 \mod 5$.
    Thus we have the trivial Dirichlet character $\chi_0$ and $5 - 1 = 4$ nontrivial Dirichlet characters mod 5, namely those given by
    \begin{align*}
        \chi_1: 1 &\mapsto 1,\ 2 \mapsto \omega_5 = \exp(2\pi i/5),\ 3 \mapsto \omega_5^3 = \exp(6\pi i/5),\ 4 \mapsto \omega_5^2 = \exp(4\pi i/5),\ \\
        \chi_2: 1 &\mapsto 1,\ 2 \mapsto \omega_5^2 = \exp(4\pi i/5),\ 3 \mapsto \omega_5 = \exp(2\pi i/5),\ 4 \mapsto \omega_5^4 = \exp(8\pi i/5),\ \\
        \chi_3: 1 &\mapsto 1,\ 2 \mapsto \omega_5^3 = \exp(6\pi i/5),\ 3 \mapsto \omega_5^4 = \exp(8\pi i/5),\ 4 \mapsto \omega_5^1 = \exp(2\pi i/5),\ \\
        \chi_4: 1 &\mapsto 1,\ 2 \mapsto \omega_5^4 = \exp(8\pi i/5),\ 3 \mapsto \omega_5^2 = \exp(4\pi i/5),\ 4 \mapsto \omega_5^3 = \exp(6\pi i/5),\ \\
    \end{align*}
\end{example}
